Article Title

### 2014-01-03 plos.tex
updated introduction

\begin{document}
% Title must be 150 characters or less
\begin{flushleft}
{\Large
Inferred partial neural correlations suggest detailed interactions in cortical circuits
}
% Insert Author names, affiliations and corresponding author email.
\\
Dimitri Yatsenko,$^{1,\ast}$, 
Kre\v{s}imir Josi\'{c}$^{2}$,
Alexander S.~Ecker$^{1,3,4}$,
Emmanouil Froudarakis$^{1}$,
R.~James Cotton$^{1}$,
Andreas S.~Tolias$^{1,5}$
\\
\bf{1} Department of Neuroscience, Baylor College of Medicine, Houston, TX, USA
\\
\bf{2} Department of Mathematics and Department of Biology and Biochemistry, University of Houston, Houston, TX, USA
\\
\bf{3}  Werner Reichardt Center for Integrative Neuroscience and Institute for Theoretical Physics, University of T\"ubingen, Germany
\\
\bf{4} Bernstein Center for Computational Neuroscience, T\"ubingen, Germany
\\
\bf{5} Department of Computational and Applied Mathematics, Rice University, Houston, TX, USA

@@

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations in neural activity measured in such recordings can reveal important aspects of the functional organization of neural circuits.  However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.  Estimation can be improved by regularization: imposing some kind of structure on the estimate.  The amount of improvement depends on how well the imposed structure accommodates the dominant dependencies in the data.  Therefore, the selection of the most efficient estimator of neural correlation matrices is an empirical question that informs about the types of dominant dependencies governing the system.

In this study, we sought most statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of partial correlations between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.  Indeed, in cross-validation tests, the covariance matrix estimator based on such approximations outperformed other regularized estimators. Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.  Because of its superior estimation performance, the partial correlation structure by these methods will likely provide a better description of the functional connectivity in dense circuits than the usual sample correlations. 
\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlations in the activity of neurons have proven useful as descriptors of the functional organization of neural circuits with implications for stimulus coding and circuit architecture.  Estimation of correlation matrices can be improved by imposing some structure on the estimate. Greatest improvements are attained when the imposed structure efficiently captures real dependencies in the data. Using fast 3D two-photon imaging of calcium signals, we recorded the activity of large and dense groups of cells in mouse visual cortex and evaluated the performance of correlation matrix estimators that imposed different kinds of structure. The correlation structure of the estimator that led to greatest improvement comprised a sparse network of partial correlations between pairs of neurons combined with partial correlations with several common latent factors. Both components were necessary for efficient estimation. Since it leads to better estimates, we proposed that this sparse+latent structure provides a better picture of the functional connectivity than that obtained from the usual sample correlation matrix in densely sampled neural recordings. As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and differences in their preferred orientations and found basic agreement with previous studies of synaptic connectivity.
\section*{Introduction}
\paragraph{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits and reflect interactions between neurons \cite{Gerstein:1964}.  This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010,Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.

However, neural correlations do not submit to ready or unambiguous mechanistic interpretation.  Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967b, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}.

\paragraph{Additional information in neural correlation matrices}
Despite these ambiguities, a complete correlation matrix provides more information than the equivalent number of isolated pairwise correlations. In early studies, the effects of correlations were extrapolated from isolated pairs to entire populations through simulation and theoretical analysis \cite{Shadlen:1998, Zohary:1994}. In contrast, modern multineuronal recordings from increasingly large and dense populations of highly interconnected neurons allow estimation of various low-dimensional representations of the entire correlation matrix, which can enable inquiry into the nature of underlying physiological mechanisms and computational principles.

For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation. Such approximations are particularly suitable for capturing shared fluctuations across the recorded population, which may arise in local circuits or be externally driven. Low-rank components can be extracted using principal component or factor analysis \cite{Chapin:1999, Peyrache:2010, Lopes:2011, Lopes:2013}.

Alternatively, the correlation matrix can be expressed through the partial pairwise correlations. The partial correlation between a pair of neurons is a measure of their linear association after the linear effects of all other neurons have been taken into account. Partial correlations carry particular significance when the relationship between the variables is approximately linear. In this case, a low pairwise partial correlation indicates that the two variables are nearly independent when conditioned on the activity of the other neurons.

Conditional independence between a pair of variables suggests a lack of direct interaction between the underlying processes. Therefore, networks of partial correlations (sometimes called \emph{association networks}) can be used to infer interactions between the components of a complex system.  This approach has been used to uncover the structure of gene interaction networks \cite{Schafer:2005, Peng:2009} and the functional connectivity between brain regions from fMRI signals \cite{Varoquaux:2012, Ryali:2012}. In multivariate normal distributions (or, more generally, \emph{elliptical distributions}) interactions are completely defined by linear effects and, therefore, partial correlations reflect conditional dependencies.  A multivariate normal distribution specified by a weighted graph of non-zero partial correlations (along with the means and variances of individual neurons) is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}. With departure from linearity, the correspondence between conditional dependence and partial correlations diminishes and can break down \cite{Loh:2012}. For example, the coupling terms in pairwise Ising models fitted to the same sample correlation matrix will produce a different network of dependencies than that inferred by the partial correlations \cite{Schneidman:2006, Tkacik:2006}.

\paragraph{Estimation of neural correlation matrices}
Here, we pursued two related aims: (a)~the improved estimation of neural correlation matrices and (b)~discovery of the low-dimensional structure of correlations in recordings of multineuronal activity. An accurate characterization of the patterns of correlations helps in interpreting them, and to better relate them to the architecture and dynamics of the underlying neuronal networks.

@@

Estimation of large covariance matrices presents a number of numerical challenges.  The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.  This mismatch leads to an increase in spurious correlations, overestimation of shared activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned estimates of the partial correlations \cite{Schafer:2005}.

Estimation can be improved through \emph{regularization}: deliberately imposing some kind of structure on the estimate. The sample estimate $C_{\sf 0}$, which is usually used to estimate covariance matrices in neuroscience, is unbiased ($\E{C_0}=\Sigma$).  However, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise.  Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases. Regularization strikes a favorable balance between bias and variability

 A regularized estimate is produced by \sq{shrinking} the unbiased estimate toward a selected low-dimensional approximation (\sq{target estimate}) of the unbiased estimate \cite{Schafer:2005, Bickel:2006}.  The selection of the optimal target estimate and the optimal amount of shrinkage can be deduced from the training data analytically or by cross-validation.  Some regularization schemes only perform target selection (\emph{dimensionality reduction, variable selection, or feature selection}) and replace the estimate completely, without shrinkage. Others only perform shrinkage toward a single target estimate (\emph{shrinkage estimators}).

Regularization can produce \emph{some} improvement even with an arbitrary target estimate (See Discussion).  Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.

Although regularized covariance estimates have become established in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization schemes for neural correlations.  Since the performance of regularization schemes depends on the dominant regularities in the data, their optimal choice depends on the type of neural recording, the temporal scale, and brain states. For example, dense recordings from local circuits with high temporal resolution in desynchronized brain states may warrant greater emphasis on partial correlations, reflecting direct interactions between neurons.  Recordings from distributed groups of cells, synchronous circuit activity, or recordings with low temporal resolution are likely to reveal common correlated activity across many cells with correlation matrices best modeled by low-rank components.  Therefore, the optimal choice of the regularization scheme, ascertained from the data by cross-validation, constitutes a statistical description of a neural circuit that informs about the dominant regularities in the data.
\paragraph{Illustration of estimation of a neural correlation matrix}
To illustrate the challenges and techniques of estimation of the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}. Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero. We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (\figref{01}{A, B}) and covariance selection (\figref{01}{C}). Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales. The close similarity of the two estimates of the correlation matrix (\figref{01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients. Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth. Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
        The outlined square fragment is magnified in {\bf B}.
        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
    }
    \label{fig:01}}
    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
\end{figure}
\paragraph{Our contribution}
Various other regularization schemes also produce improved estimates.  In this study, we compared the sample covariance matrix and four regularized estimators based on distinct correlation structures as follows:
\begin{description}
\item[$C_{\sf 0}$] -- sample covariance matrix, the usual estimator used in previous studies.
\item[$C_{\sf diag}$] -- linear shrinkage toward a diagonal covariance matrix.
\item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis with an optimal number of factors.
\item[$C_{\sf sparse}$] -- sparsified precision matrix,  \emph{covariance selection} \cite{Dempster:1972, Friedman:2008}, without shrinkage.   
\item[$C_{\sf sparse+latent}$] -- precision matrix composed as the sum of a sparse and a low-rank component \cite{Chandrasekaran:2010, Ma:2013}.
\end{description}

We also performed a cross-validated evaluation to establish the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex. We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.  This estimator produced a sparse network of partial correlations between the observed neurons combined with correlations between several inferred latent units. Since it led to better estimates, we proposed that this structure provided a more physiologically relevant representation of the circuit’s functional connectivity than that obtained from the sample correlation matrix.  We showed that the partial correlation structure had stronger dependence on physical distances and differences in orientation preference than sample correlations. In Discussion, we outline possibly approaches to corroborate the relationship between the structure of the functional connectivity revealed by this and similar method to the anatomical organization of the circuit.

@@

The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between a sparse set of pairs of observed cells (\figref{02}{C}).  This assumption is enforced by setting to zero an optimal subset of the off-diagonal elements of the precision matrix.  Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (\figref{02}{C}). The estimator takes the form
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1},
\end{equation}
where $S$ is a sparse matrix with a large fraction of zeros off the diagonal. The estimate has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) of $S$.

@@

Then the cross-validation loss is in fact an unbiased estimate of the regular loss.
\begin{equation*}
    \E[C]{\E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}}=\E[C]{\loss{C,\E[C_{\sf 0}]{C_{\sf 0}^\prime}}}=\E[C]{\loss{C,\Sigma}}
\end{equation*}

@@

\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf The sparse+latent estimator $C_{\sf sparse+latent}$ outperforms the other estimators on neural data.}
    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+latent}$. 
    The histograms are based on 31 imaged sites in 24 mice. 
    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
    The arrow heads indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}
{\bf A} $C_{\sf sparse+latent}$ outperforms $C_{\sf 0}$: median improvement 0.16 nats/neuron, $p=3.9\times 10^{-5}$.
    {\bf B} $C_{\sf sparse+latent}$ outperforms $C_{\sf diag}$: median improvement 0.0032 nats/neuron, $p=2.2\times 10^{-3}$.
    {\bf C} $C_{\sf sparse+latent}$ outperforms $C_{\sf factor}$: median improvement 0.16 nats/neuron, $p=6.9\times 10^{-5}$.
    {\bf D} $C_{\sf sparse+latent}$ outperforms $C_{\sf sparse}$: median improvement 0.0016 nats/neuron, $p=5.5\times 10^{-6}$.
} \label{fig:04}}
    {\includegraphics{./figures/Figure04.pdf}}
\end{figure}

\paragraph{Correlation structure and circuit architecture}
Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the low-dimensional structure revealed by this estimator for individual imaged sites. By design, the sparse+latent estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units. If this estimate dominates, we can hypothesize that it is better reveals underlying physiological interactions than the usual correlations. In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.

@@

\section*{Discussion}
\paragraph{Functional connectivity}
How the anatomical organization of neural circuits gives rise to their function is a foremost question in neuroscience.  Effective phenomenological descriptions of neural activity, once discovered, enable investigation of underlying physiological mechanisms.  For example, the discoveries of receptive fields and tuning properties of individual cells, over the past 60 years, have led to fruitful hypotheses about the underlying patterns of synaptic connectivity, principles of neural coding, and to the discovery of other organizational principles such as various cortical feature maps \cite{Reid:2012}.

>>>>

\begin{document}
% Title must be 150 characters or less
\begin{flushleft}
{\Large
Partial neural correlations suggest detailed interactions in cortical circuits
}
% Insert Author names, affiliations and corresponding author email.
\\
Dimitri Yatsenko,$^{1,\ast}$, 
Kre\v{s}imir Josi\'{c}$^{2}$,
Alexander S.~Ecker$^{1,3,4}$,
Emmanouil Froudarakis$^{1}$,
R.~James Cotton$^{1}$,
Andreas S.~Tolias$^{1,5}$
\\
\bf{1} Department of Neuroscience, Baylor College of Medicine, Houston, TX, USA
\\
\bf{2} Department of Mathematics and Department of Biology and Biochemistry, University of Houston, Houston, TX, USA
\\
\bf{3}  Werner Reichardt Center for Integrative Neuroscience and Institute for Theoretical Physics, University of T\"ubingen, Germany
\\
\bf{4} Bernstein Center for Computational Neuroscience, T\"ubingen, Germany
\\
\bf{5} Department of Computational and Applied Mathematics, Rice University, Houston, TX, USA

@@

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations in neural activity measured in such recordings can reveal important aspects of the functional organization of neural circuits.  However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.  Estimation can be improved by regularization, \emph{i.e.}\;by imposing a structure on the estimate.  The amount of improvement depends on how closely the assumed structure represents dependencies in the data. Therefore, the selection of the most efficient correlation matrix estimator for a given neural circuit must be determined empirically.  Furthermore, the identity and structure of the most efficient estimator informs about the types of dominant dependencies governing the system.

We sought the most statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix should be most efficiently represented as the combination of a sparse matrix of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.  Indeed, in cross-validation tests, the covariance matrix estimator with such structure consistently outperformed other regularized estimators. The connectivity patterns inferred from the sparse component of the estimate manifested a strong relationship to the geometrical arrangement and orientation tuning properties of cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distance and difference in orientation preference whereas the negative \sq{inhibitory} connections were less selective.  Because of its superior performance, this estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than sample correlations.

\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlations in the activity of neural populations have proven useful to describe their functional connectivity.  Correlations may arise from various underlying processes that reflect differently on the structure of the correlation matrix:  Common network fluctuations or external inputs are well approximated by low-rank representations whereas linear effects between specific pairs of neurons are well approximated by fitting their pairwise partial correlations.  What kinds of correlation effects dominate the population activity in cortical microcircuits? To answer this question, we imposed various correlation structures on empirical estimates of neural correlation matrices and evaluated which of them produced greatest estimation improvement.  In fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the greatest improvement was produced by decomposing the correlation matrix as a sparse network of partial correlations and a low-rank component.  Since it leads to the best estimates, we propose that this structure provides a better picture of the functional connectivity than that obtained from the usual correlations. As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and to differences in their preferred orientations and found basic resemblance to results from studies of synaptic connectivity.
\section*{Introduction}
\paragraph{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}\;the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits and reflect interactions between neurons \cite{Gerstein:1964}.  This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010,Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.

However, neural correlations do not submit to ready or unambiguous mechanistic interpretation.  Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}.

\paragraph{Structure  of correlation matrices}
Modern multineuronal recordings allow measuring entire correlation matrices from increasingly large and dense populations of highly interconnected neurons. Multivariate analysis techniques based on large correlation matrices provide information that could not be extracted from an equivalent number of isolated measurements of pairwise correlations.  For example, when the correlation matrix is approximated in a form with a reduced number of variables, this \emph{reduced representation} can help isolate various aspects of the functional connectivity in neural circuits and possibly point to underlying physiological mechanisms and computational principles.

One reduced representation of the correlation matrix is its low-rank approximation.  Low-rank approximations are particularly suitable for capturing common fluctuations or common external inputs into the circuit and can be extract using principal component analysis, factor analysis, or  \cite{Chapin:1999, Peyrache:2010, Lopes:2011, Lopes:2013}.

Another reduced representation of the correlation matrix can be obtained by \emph{covariance selection}, \emph{i.e.}\;by selecting a subset of the pairwise \emph{partial correlation} that are sufficient to reproduce the entire correlation matrix \cite{Dempster:1972, Friedman:2008}. The partial correlation between a pair of neurons is a measure of their linear association after the effects of the rest of the network have been taken into account \cite{Anderson:2003}. Partial correlations are particularly meaningful when relationships between variables are mostly linear and large fraction of the variables are observed.  In this case, partial correlations approximate conditional dependency.  A lack of conditional dependency between two neurons suggests a lack of direct interaction between them. Networks of partial correlations (sometimes called \emph{association networks}) have been used to infer gene interaction networks \cite{Schafer:2005, Peng:2009} and interactions between brain regions in brain imaging studies \cite{Varoquaux:2012, Ryali:2012}.

\paragraph{Improved estimation performance by regularization}
Imposing some structure on an estimate of the covariance matrix can improve its estimation performance.

Estimation of covariance matrices from large populations presents a number of numerical challenges:  The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.  This mismatch leads to an increase in spurious correlations, overestimation of common activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned estimation of the partial correlations \cite{Schafer:2005}.  In neuroscience studies, the true correlation matrix is usually estimated by the \emph{sample correlation matrix} and these two concepts are rarely distinguished.  The sample correlation matrix is unbiased but, because it has many free parameters, it is highly sensitive to sampling noise. As a result, on average, the sample correlation matrix is far from the true correlation matrix.  In contrast, reduced estimates are typically less susceptible to sampling noise although can introduce their respective biases.

In statistics, the technique of deliberately imposing some structure on an estimate in order to improve its performance is known as \emph{regularization} \cite{Schafer:2005, Bickel:2006}.    To \sq{impose structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation known as the shrinkage target.  The selection of the optimal target estimate and the optimal amount of shrinkage can be deduced from the training data either analytically or by cross-validation.  Some regularization schemes only perform target selection (\emph{dimensionality reduction, variable selection, or feature selection}) and replace the estimate completely, without shrinkage. Others only perform shrinkage toward a single target estimate (\emph{shrinkage estimators}).  Yet others can do both target selection and shrinkage toward the target.

Although regularized covariance estimates have become commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.  
\paragraph{Illustration of estimation of a neural correlation matrix}
To illustrate the challenges and techniques of estimation of the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972, Friedman:2008}. Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix (inverse of the covariance matrix) while setting the rest to zero.  Zeros in the off-diagonal elements of the precision matrix indicate zero partial correlations between the corresponding pair of neurons.   Therefore, covariance selection amounts to finding an approximation of the covariance matrix so that a large fraction of neuron pairs are linearly independent of each other when conditioned on the activity of all the other observed cells.

We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (\figref{01}{A, B}) and covariance selection (\figref{01}{C}). Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales. The close similarity of the two estimates of the correlation matrix (\figref{01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients.

Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth. Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
        The outlined square fragment is magnified in {\bf B}.
        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
    }
    \label{fig:01}}
    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
\end{figure}
\paragraph{Conceptual questions and basic approach}
What structures of correlation matrices are most suitable for the description of the multineuronal activity in specific neural circuits and brain states? For example, are correlations in visual cortex under during visual stimulation best explained by common fluctuations or by local interactions within the recorded microcircuit?

We pursued two related aims: (a)~the improved estimation of neural correlation matrices and (b)~discovery of the low-dimensional structure of correlations in recordings of multineuronal activity. An accurate characterization of the patterns of correlations helps in interpreting them, and to better relate them to the architecture and dynamics of the underlying neuronal networks.

 In this study, we compared the sample covariance matrix and four regularized estimators based on distinct correlation structures:
\begin{description}
\item[$C_{\sf 0}$] -- sample covariance matrix, the usual estimator used in previous studies.
\item[$C_{\sf diag}$] -- linear shrinkage toward a diagonal covariance matrix.
\item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis with an optimal number of factors.
\item[$C_{\sf sparse}$] -- sparsified precision matrix,  \emph{covariance selection} \cite{Dempster:1972, Friedman:2008}, without shrinkage.   
\item[$C_{\sf sparse+latent}$] -- precision matrix composed as the sum of a sparse and a low-rank component \cite{Chandrasekaran:2010, Ma:2013}.
\end{description}

We also performed a cross-validated evaluation to establish the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex. We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.  This estimator produced a sparse network of partial correlations between the observed neurons combined with correlations between several inferred latent units. Since it led to better estimates, we propose that this structure provides a more physiologically relevant representation of the circuit’s functional connectivity than that obtained from the sample correlation matrix.  We showed that the partial correlation structure had stronger dependence on physical distances and differences in orientation preference than sample correlations. In Discussion, we outline possibly approaches to corroborate the relationship between the structure of the functional connectivity revealed by this and similar method to the anatomical organization of the circuit.

@@

Various other regularization schemes also produce improved estimates

@@

The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between pairs of observed cells (\figref{02}{C}).  This assumption is enforced by fitting only a subset of the off-diagonal elements of the precision matrix while setting the rest to zero.    Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (\figref{02}{C}). The estimator takes the form
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1},
\end{equation}
where $S$ is a sparse matrix with a large fraction of zeros off the diagonal. The estimate has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) of $S$.

@@

Then the cross-validation loss is in fact an unbiased estimate of loss.
\begin{equation*}
    \E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}=\loss{C,\E[C_{\sf 0}]{C_{\sf 0}^\prime}}=\loss{C,\Sigma}
\end{equation*}

@@

\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf The sparse+latent estimator $C_{\sf sparse+latent}$ outperforms the other estimators on neural data.}
    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+latent}$. 
    The histograms are based on 31 imaged sites in 24 mice. 
    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
    The arrow heads indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}
\\{\bf A.} $C_{\sf sparse+latent}$ outperforms $C_{\sf 0}$: median improvement 0.16 nats/neuron, $p=3.9\times 10^{-5}$.
\\{\bf B.} $C_{\sf sparse+latent}$ outperforms $C_{\sf diag}$: median improvement 0.0032 nats/neuron, $p=2.2\times 10^{-3}$.
\\{\bf C.} $C_{\sf sparse+latent}$ outperforms $C_{\sf factor}$: median improvement 0.16 nats/neuron, $p=6.9\times 10^{-5}$.
\\{\bf D.} $C_{\sf sparse+latent}$ outperforms $C_{\sf sparse}$: median improvement 0.0016 nats/neuron, $p=5.5\times 10^{-6}$.
} \label{fig:04}}
    {\includegraphics{./figures/Figure04.pdf}}
\end{figure}

\paragraph{Correlation structure and circuit architecture}
Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the low-dimensional structure revealed by this estimator for individual imaged sites. By design, the sparse+latent estimator finds the mixture of sparse partial correlations between visible neurons and common fluctuations described as latent units. If this estimate dominates, we can hypothesize that it better describes underlying physiological interactions than other estimators. In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.

@@

\section*{Discussion}
\paragraph{Functional connectivity}
How the anatomical organization of neural circuits gives rise to their function is a foremost question in neuroscience.  Effective phenomenological descriptions of neural activity, once discovered, enable investigation of underlying physiological mechanisms.  For example, the discoveries of receptive fields and tuning properties of single neurons, over the past 60 years, have led to fruitful hypotheses about the underlying patterns of synaptic connectivity, principles of neural coding, and to the discovery of other organizational principles such as various cortical feature maps \cite{Reid:2012}.
